A new method to estimate dark matter halo concentrations

### 2015-01-13 paper/article.tex
writing the new method section

The Navarro-Frenk-White density profile can be written as

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2}, 
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is known as the scale radius, the radius that marks the transition
between the two power law behaviour in the $\rho\propto r^{-1}$ for
$r<r_s$ and $\rho\propto r^{-3}$ for  $r>r_s$.

We define the virial radius of a halo, $r_v$, as the boundary of the
spherical volume that encloses an average density of $\Delta_h$ times
the average density of the Universe. The corresponding mass $M_{v}$,
the virial mass, can be written as $M_{v} =
\frac{4\pi}{3}\bar{\rho}\Delta_h r_v^3$.

@@

We can now express the same quantity in terms of dimensionless the
variables $x\equiv r/r_v$ and $m\equiv M(<r)/M_v$,
%
\begin{equation}
m(<x) =
\frac{1}{A}\left[\ln\left(1+xc\right)-\left(\frac{xc}{xc+1}\right)\right],
\label{eq:profile}
\end{equation}
%
where 
%
\begin{equation}
A=\left[\ln\left(1+c\right)-\left(\frac{c}{c+1}\right)\right],
\end{equation}
%
and the parameter $c$ is known as the concentration $c\equiv r_v/r_s$.

@@

It is also customary to express the mass of the halo inte terms of the
circular velocity $V_{c}=\sqrt{GM(<r)/r}$. From this we can
define a new dimensionless circular velocity $v(<x)\equiv
V_{c}(<r)/V_{c}(<r_v)$, using the result in Equation \ref{eq:profile}
to have:

\begin{equation}
v(<x)=\sqrt{\frac{1}{A}\left[\frac{\ln\left(1+xc\right)}{x}-\frac{c}{xc+1}\right]},
\end{equation}

@@

\section{A new Approach to Halo Fitting}
\label{sec:method}

>>>>

\subsection{Density profile}

The NFW density profile can be written as

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2}, 
\label{eq:definition}
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is known as the scale radius, the radius that marks the transition
between the power law scaling $\rho\propto r^{-1}$ for
$r<r_s$ and $\rho\propto r^{-3}$ for  $r>r_s$.

We define the virial radius of a halo, $r_v$, as the boundary of the
spherical volume that encloses a density of $\Delta_h$ times
the average density of the Universe. The corresponding mass $M_{v}$,
the virial mass, can be written as $M_{v} =
\frac{4\pi}{3}\bar{\rho}\Delta_h r_v^3$.

@@

We express the same quantity in terms of dimensionless
variables $x\equiv r/r_v$ and $m\equiv M(<r)/M_v$,
%
\begin{equation}
m(<x) =
\frac{1}{A}\left[\ln\left(1+xc\right)-\left(\frac{xc}{xc+1}\right)\right],
\label{eq:profile}
\end{equation}
%
where 
%
\begin{equation}
A=\ln\left(1+c\right)-\left(\frac{c}{c+1}\right),
\end{equation}
%
and the parameter $c$ is known as the concentration $c\equiv r_v/r_s$.

From this normalization value and for later convenience we define the
following function 
%
\begin{equation}
f(x) = \ln\left(1+x\right)-\left(\frac{x}{x+1}\right).
\label{eq:f_NFW}
\end{equation}
%

@@

It is also customary to express the mass of the halo in terms of the
circular velocity $V_{c}=\sqrt{GM(<r)/r}$. From this we can
define a new dimensionless circular velocity $v(<x)\equiv
V_{c}(<r)/V_{c}(<r_v)$, using the result in Eq. \ref{eq:profile}
to have:

\begin{equation}
v(<x)=\sqrt{\frac{1}{A}\left[\frac{\ln\left(1+xc\right)}{x}-\frac{c}{xc+1}\right]},
\end{equation}
%
this normalized profile always shows a maximum provided that the
concentration is larger than a values of $c>???$. 
It is possible to show that for the NFW profile the maximum is
provided by

\begin{equation}
\mathrm{max}(v(<x)) = \sqrt{\frac{c}{x_{\rm max}}\frac{f(x_{\rm
      max})}{f(c)}}, 
\label{eq:max_v}
\end{equation}
where $x_{\rm max}=2.163$ and the function $f(x)$ was defined in Eq. (\ref{eq:f_NFW}).

@@

\section{A new approach to estimate the halo concentration}
\label{sec:method}

Once the density profile is expressed in dimensionless variables the
only free variable is the concentration. 
There are two main methods to estimate this values in dark matter
halos extracted from N-body simulations.
The first method bins the particles in the logarithm of the radial
coordinate from the halo center to estimate the density in each
logarithmic bin. 
Then a direct fit to the density as a function of the radial
coordinate is performed. This method has been most recently advocated
by \cite{Ludlow2014} to study the mass-concentration-redshift relation
of dark matter halos using the Millennium Simulation Series.

A second method uses the circular velocity profile. 
As it was shown in Fig. (\ref{fig:velocity}) the circular velocity
shows a maximum for all profiles with concentration values larger than
$c>???$. 
For the numerical halos wne can then find the value of $x$ for which
the normalized circular velocity $v(<x)$ shows a maximum and solve
numerically for the corresponding value of the concentration using
Eq. (\ref{eq:max_v}).  This method has been used by \cite{Klypin2014}
to study the mass-concentration-redshift relation using the Multidark
Simulation Suite.
